\bf A Hierarchical Model for Heterogenous Reliability Field Data

### 2018-06-09 paper/m-lb-m-technometrics.tex
Fixes 2 small typos in May 31 version
Incorporates Meeker edits from June 8

Following the suggestions given in \cite{LiMeeker2014} and Section 15.2 of \cite{intervals} we use an alternative parameterization
where the usual scale parameter $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$ (which is also a scale parameter). Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0
\end{align*}
There are three important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}
 \cite{gelman2004parameterization} provides a more general discussion of reparameterization and Bayesian inference.

@@

\subsection{Likelihood}
\label{subsec:likehood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time, the age of unit $i$ when data reporting commenced. Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of the parameters $\bm{\theta} = (\pi,t_{p_1}, \sigma_1, t_{p_2}, \sigma_2)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is
\begin{equation}
\label{equation:likelihood}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation}

@@

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the traditional parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions (See the Supplementary Material for standard density plots of the priors).  Using this convention the prior distributions used in our analyses are
%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{p_1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
t_{p_2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
$$\pi \sim \op{Logit-normal}\langle 0.001,\, 0.71 \rangle,\quad
\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle$$
where $p_1=0.2$ and $p_2=0.5$.
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

@@

For all of the nested models, the parameters for the early failure mode are common across sub-populations.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of the $\pi_g$ across the subpopulations.  Also the lifetime distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest.

\subsection{Model selection}
The use of Bayes factors is a classical apporach to compare alternative model specifications.  However, they can be problematic, especially in high dimensional settings where there is no concensus on how to properly compute the required marginal likelihood \citep{robert2016}. Moreover, from the point of view of a consumer or purchaser, the goal of an analysis such as this is to rate manufacturers and/or to inform future purchases based on predicted product performance. Therefore, we should compare model specifications based on accurate prediction for future units of previously observed products.  Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
posterior draws from a fitted model.  Following \cite{vehtari}, the log pointwise predictive density (lpd), is one method to quantify predictive accuracy,

@@

\paragraph{Model 2} $t_{p_{2}g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the early failure mode:
$$ \pi \sim \op{Logit-normal}\langle 0.007,\, 0.26 \rangle,$$
$$\sigma_1 \sim \op{Lognormal}\langle 0.14,\, 7.1\rangle,$$ 
$$t_{p_1} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle.$$
where $p_{1}=0.5$.

@@

Where applicable, prior distributions on hyperparameters are as listed below. We use half-Cauchy prior distributions on hierarchical scale parameters. As for the location hyperparameters, we choose weakly informative prior distributions centered around the corresponding prior mean for the non-hierarchical model.
$$\eta_{\pi} \sim \op{Normal}(-3, 1^2),\quad \tau_{\pi} \sim \op{Half-Cauchy}(0, 1)$$
$$\eta_{\sigma_2}  \sim \op{Normal}(0, 2^2),\quad
\tau_{\sigma_2}  \sim \op{Half-Cauchy}(0, 1)$$
$$\eta_{t_{p_2}}  \sim \op{Normal}(9, 2^2),\quad
\tau_{t_{p_2}}  \sim \op{Half-Cauchy}(0, 1)$$
where$p_{2}=0.2$.

@@

We can also compare posterior estimates of time to failure under Models 1-4 for each drive-model individually.  For these comparisons we calculate pointwise the posterior median of the proportion failing over a grid. That is, for every model, and for a fixed set, $\tilde{t}=\{\tilde{t}_1,\ldots,\tilde{t}_M\}$, we compute 
\begin{equation}
\label{pointwise-medians}
\op{median}\left\{\op{H}\left(\tilde{t}_m|\pi_{g}^{(s)},t_{p_{1}g}^{(s)},\sigma_{1g}^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
\end{equation}
where $\theta^{(s)}$ is posterior sample $s$ from $p(\theta|y)$.
%Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.

@@

%stopped here
The GLFP model fits for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (See the Supplementary Material for similar plots for the other drive-models).  The step functions again correspond to the adjusted K-M estimates.  As model complexity increases, the hierarchical-model GLFP curves become more flexible and are better able to describe the observed failure data.  All of the GLFP models appear to produce higher estimates of the proportion failing, relative to K-M, for Drive-model 2. This behavior is not unexpected, however, because hierarchical models improve model stability through shrinkage; in this context, shrinkage is toward an average cdf. Model 4 is in closest agreement.

@@

For most drive-models, the estimates based on the observed data look similar to those based on the replicated data.  There are a few drive-models, however where this is not the case. Drive-model 14 shows a discrepancy between the GFLP fit and the K-M fit in the right tail. In particular, the GLFP predictions appear conservative, overpredicting the proportion failing after about 2.5 years. Due to the large number of failures for this drive-model, the posterior for the hierarchical model is similar to the stand-alone model. Other drive-models showing a lack of fit are 4 and 34. We first address Drive-model 4; our diagnosis of 34 is analogous, so the following comments extend to that drive-model as well.

@@

Consumers of manufactured products usually lack the data, expertise, or motivation to perform a failure analysis to determine the cause of failure.  Their interest is primarily in the lifetime distribution of a product rather than the specific cause of failure.  The ability to fit marginal models, such as GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product sub-populations.  Because the Bayesian approach provides full posterior distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the actual application.

@@

\item[Appendices, stan model code and plots] Description of truncation adjustment to Kaplan Meier estimator, global average, Stan code for Model 4 as well as the complete set of plots similar to Figures 5 and 6 (supplementary.pdf)

>>>>

Following the suggestions given in \cite{LiMeeker2014} and Section 15.2 of \cite{intervals} we use an alternative parameterization
where the usual scale parameter $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$ (which is also a scale parameter). Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
\end{align*}
There are three important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}
 \cite{gelman2004parameterization} provides a more general discussion of reparameterization and Bayesian inference.

@@

\subsection{Likelihood}
\label{subsec:likehood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time, the age of unit $i$ when data reporting commenced. Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of the parameters $\bm{\theta} = (\pi,t_{p_1}, \sigma_1, t_{p_2}, \sigma_2)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is
\begin{equation}
\label{equation:likelihood}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}.
\end{equation}

@@

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the traditional parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions (See the Supplementary Material for standard density plots of the prior distributions).  Using this convention the prior distributions used in our analyses are
%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{p_1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
t_{p_2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
$$\pi \sim \op{Logit-normal}\langle 0.001,\, 0.71 \rangle,\quad
\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle$$
where $p_1=0.2$ and $p_2=0.5$.
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

@@

For all of the nested models, the parameters for the early failure mode are common across sub-populations.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for early failures provides a meaningful interpretation and comparison of the $\pi_g$ across the subpopulations.  Also the lifetime distribution of the defective subpopulation (as opposed to the proportion defective) is not of high interest in practical applications.

\subsection{Model selection}
The use of Bayes factors is a classical apporach to compare alternative model specifications.  However, they can be problematic, especially in high dimensional settings where there is no concensus on how to properly compute the required marginal likelihood \citep{robert2016}. Moreover, from the point of view of a consumer or purchaser, the goal of an analysis such as this is to rate manufacturers and/or to inform future purchases based on predicted product performance. Therefore, we should compare model specifications based on accurate prediction for future units of previously observed products.  Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
posterior draws from a fitted model.  Following \cite{vehtari}, the log pointwise predictive density (lpd) is one method to quantify predictive accuracy,

@@

\paragraph{Model 2} $t_{p_{2}g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the early failure mode:
$$ \pi \sim \op{Logit-normal}\langle 0.007,\, 0.26 \rangle,$$
$$\sigma_1 \sim \op{Lognormal}\langle 0.14,\, 7.1\rangle,$$ 
$$t_{p_1} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle,$$
where $p_{1}=0.5$.

@@

Where applicable, prior distributions on hyperparameters are as listed below. We use half-Cauchy prior distributions on hierarchical scale parameters. As for the location hyperparameters, we choose weakly informative prior distributions centered around the corresponding prior mean for the non-hierarchical model.
$$\eta_{\pi} \sim \op{Normal}(-3, 1^2),\quad \tau_{\pi} \sim \op{Half-Cauchy}(0, 1)$$
$$\eta_{\sigma_2}  \sim \op{Normal}(0, 2^2),\quad
\tau_{\sigma_2}  \sim \op{Half-Cauchy}(0, 1)$$
$$\eta_{t_{p_2}}  \sim \op{Normal}(9, 2^2),\quad
\tau_{t_{p_2}}  \sim \op{Half-Cauchy}(0, 1)$$
where $p_{2}=0.2$.

@@

We can also compare posterior estimates of time to failure under Models 1-4 for each drive-model individually.  For these comparisons we calculate pointwise the posterior median of the proportion failing over a grid. That is, for every model, and for a fixed set, $\tilde{t}=\{\tilde{t}_1,\ldots,\tilde{t}_M\}$, we compute 
\begin{equation}
\label{pointwise-medians}
\op{median}\left\{\op{H}\left(\tilde{t}_m|\pi_{g}^{(s)},t_{p_{1}g}^{(s)},\sigma_{1g}^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
\end{equation}
where $\theta^{(s)}$ is a posterior sample $s$ from $p(\theta|y)$.
%Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.

@@

%stopped here
The GLFP model fits for a representative selection of drive-models (Drive-models 2, 9, 14, and 40) are presented in Figure \ref{fig:mod_comp_leg} (see the Supplementary Material for similar plots for the other drive-models).  The step functions again correspond to the adjusted K-M estimates.  As model complexity increases, the hierarchical-model GLFP curves become more flexible and are better able to describe the observed failure data.  All of the GLFP models appear to produce higher estimates of the proportion failing, relative to K-M, for Drive-model 2. This behavior is not unexpected, however, because hierarchical models improve model stability through shrinkage; in this context, shrinkage is toward an average cdf. Model 4 is in closest agreement.

@@

For most drive-models, the estimates based on the observed data look similar to those based on the replicated data.  There are a few drive-models, however where this is not the case. Drive-model 14 shows a discrepancy between the GFLP fit and the K-M fit in the right tail. In particular, the GLFP predictions overpredict the proportion failing after about 2.5 years. Due to the large number of failures for this drive-model, the posterior for the hierarchical model is similar to the stand-alone model. Other drive-models showing a lack of fit are 4 and 34. We first address Drive-model 4; our diagnosis of 34 is analogous, so the following comments extend to that drive-model as well.

@@

Consumers of manufactured products usually lack the data, expertise, or motivation to perform a failure analysis to determine the cause of failure.  Their interest is primarily in the lifetime distribution of a product rather than the specific cause of failure.  The ability to fit marginal models, such as the GLFP, allow consumers and manufacturers to accurately model products with bathtub hazard lifetime distributions and make comparisons across different product sub-populations.  Because the Bayesian approach provides full posterior distributions once a model is fit, the analyst can easily estimate a functional, such as a quantile or depreciation factor, that allow products to be compared using statistics that are meaningful in the context of the actual application.

@@

\item[Appendices, stan model code, plots, and tables] Description of truncation adjustment to Kaplan Meier estimator global average; Stan code for Model 4; complete set of plots similar to Figures 5 and 6; plots of prior distributions for Models 1-4; sensitivity analysis of the prior distributions for Model 4;
and a complete list of brand names corresponding to the model numbers used in the analysis (supplementary.pdf)
